Gamma-Ray Burst Afterglows In The Multi-Messenger Era: Numerical Models and Closure Relations

### 2019-09-19 structuredJets.tex
Added closure plot!

The $g$ parameter evolves with time as $\theta^*$ sweeps from the jet edge to the core.  This produces deviations in the light curve from a pure power law.  However, we find ultimately these deviations are not too large and the average slope is well approximated by $\geff  = g(\theff)$, where $\theff$ is a fiducial angular section which depends on viewing angle and potentially other structure parameters.  For the Gaussian model $\theff$ is simply half the viewing angle, for the power law model a suitable fitting function was found through comparison to a grid of \afterglowpy{} light curves:
\begin{align}
	\theff &= \frac{\thobs}{2}&& \text{Gaussian} \\
	\theff &= \thobs \big [1.8+2.1b^{-1.25}  && \text{power law} \nonumber  \\
	 	& \qquad + (0.49-0.86b^{-1.15}) \thobs/\thC \big ] ^{-1/2} &&\label{eq:theff}
\end{align}
In the expressions for $\geff$ it is reasonable to assume $2 \tan((\thobs-\theff)/2) \approx \thobs-\theff$ leading to the manageable expressions:
\begin{align}
	\geff &= \frac{\thobs^2}{4\thC^2} && \text{Gaussian}\ , \\
	\geff &= \frac{2 b (\thobs-\theff)\theff}{b \thC^2+\theff^2} && \text{power law}\ . \label{eq:geff}
\end{align}
These values may be used in Table \ref{tab:slopes} for an approximate power law model of a structured jet with an appropriate viewing angle dependent temporal evolution.  These analytic results are in agreement with previous work on structured jets \citep{Rossi:2004aa} and the trans-relativistic numerical model described in Section \ref{sec:numerical}.

In the structured phase, the temporal slope depends directly on the ratio $\thobs/\thC$.  Therefore any measurement of the effective power law slope of an afterglow light curve in the structured phase can be used to measure this parameter.

@@

Figure \ref{fig:slopesG} shows the structured phase power law temporal slope as a function of $g$ for spectral regimes $D$--$H$. In the $G$ and $H$ regimes the spectral slope is a function of $p$, this dependence carries over into the temporal slope as well. The temporal slope is an increasing function of $g$ in all regimes.  At sufficiently large $g$ all regimes exhibit rising light curves, the specific $g$ at which the light curve begins to rise is regime dependent.

>>>>

@@

Figure \ref{fig:slopesG} shows the structured phase power law temporal slope $\alpha_{\mathrm{struct}}$ as a function of $g$ for spectral regimes $D$--$H$. In the $G$ and $H$ regimes the spectral slope $\beta$ is a function of $p$, this dependence carries over into the temporal slope as well. The temporal slope $\alpha_{\mathrm{struct}}$ is an increasing function of $g$ in all regimes.  At sufficiently large $g$ all regimes exhibit rising light curves, the specific $g$ at which the light curve begins to rise is regime dependent.

The $g$ parameter evolves with time as $\theta^*$ sweeps from the jet edge to the core.  This produces deviations in the light curve from a pure power law.  However, we find ultimately these deviations are not too large and the average slope is well approximated by $\geff  = g(\theff)$, where $\theff$ is a fiducial angular section which depends on viewing angle and potentially other structure parameters.  For the Gaussian model $\theff$ is simply half the viewing angle, for the power law model a suitable fitting function was found through comparison to a grid of \afterglowpy{} light curves:
\begin{align}
	\theff &= \frac{\thobs}{2}&& \text{Gaussian} \\
	\theff &= \thobs \big [1.8+2.1b^{-1.25}  && \text{power law} \nonumber  \\
	 	& \qquad + (0.49-0.86b^{-1.15}) \thobs/\thC \big ] ^{-1/2} &&\label{eq:theff}
\end{align}
In the expressions for $\geff$ it is reasonable to assume $2 \tan((\thobs-\theff)/2) \approx \thobs-\theff$ leading to the manageable expressions:
\begin{align}
	\geff &= \frac{\thobs^2}{4\thC^2} && \text{Gaussian}\ , \\
	\geff &= \frac{2 b (\thobs-\theff)\theff}{b \thC^2+\theff^2} && \text{power law}\ . \label{eq:geff}
\end{align}
These values may be used in Table \ref{tab:slopes} for an approximate power law model of a structured jet with an appropriate viewing angle dependent temporal evolution.

\begin{figure*}
	\includegraphics[width=\textwidth]{figs/lc_closure_multi.pdf}
	\caption{Optical afterglow light curves in the \emph{structured} phase with corresponding closure relation slopes $\alpha_{\mathrm{struct}}$.  Each panel shows light curves with $\thobs/\thC$ = 2, 4, 6, and 8 (solid lines in red, green, orange, and blue respectively) calculated with \afterglowpy{} and corresponding power law segments $F_\nu \propto \tobs^{\alpha_{\mathrm{struct}}}$ calculated from the closure relations in Table \ref{tab:slopes} and the $\geff$ values from Equation \eqref{eq:geff}.  The left panel shows a Gaussian jet, the middle a $b=2$ power law jet, and the right a $b=6$ power law jet.  The closure relations $\alpha_{\mathrm{struct}}$ capture the approximate power law slope of the numerical light curves, with the agreement improving for larger values of $\thobs/\thC$. The light curves use a fixed $\thobs$ = 0.3 rad, $\thW = 0.25$ rad, and vary $\thC$.  Other parameters are fixed: $E_0 = 10^{53}$ erg, $n_0=10^{-3}$ cm$^{-3}$, $p=2.2$, $\epse = 10^{-1}$, $\epsB = 10^{-3}$, $d_L=10^{28}$ cm and $z=0.5454$. \label{fig:closure}}
\end{figure*}

Figure \ref{fig:closure} shows \afterglowpy{} light curves for Gaussian, $b=2$ power law, and $b=6$ power law jets at various values of $\thobs / \thC$ and the corresponding $\alpha_{\mathrm{struct}}$ slopes calculated using $\geff$.  The analytic closure relations capture the light curve behaviour in the structured phase for all 3 models, with somewhat better agreement at larger values of $\thobs/\thC$.  The Gaussian jet, with its strong energy gradient, is most sensitive to changes in $\thobs/\thC$.  The $b=6$ power law, with its more energetic wings, displays less sensitivity to changes in $\thobs/\thC$.  The $b=2$ power law displays almost no variation at all in the temporal decay rate.  The overall behaviour of the Gaussian and $b=2$ power law jets are in agreement with \citet{Rossi:2004aa} (R04, see for comparison their Figure 12 and 16).  Their $\thobs/\thC = 4$ Gaussian jet showed a mildly decaying slope instead of the rising light curve seen in Figure \ref{fig:closure}, due to the choice of different values of $p$.  If $\thobs/\thC=4$, then a Gaussian jet will have also $\geff = 4$ by Equation \eqref{eq:geff}, and Figure \ref{fig:slopesG} shows $\alpha_{\mathrm{struct}}$ will be negative for $p=2.5$ in the G regime as in R04.

In the structured phase, the temporal slope depends directly on $\geff$ in a way independent of the particular jet model.  If a particular structure is assumed, then $\thobs/\thC$ may easily be inferred from $\geff$, and hence from $\alpha_{\mathrm{struct}}$.
